---
title: "coursework2"
output: html_document
---

```{r}
library(tidyverse)
library(zoo)
library(tseries)
library(knitr)
library(xts)
library(lubridate)
library(broom)
```

## Connect to WRDS API
```{r}
library(RPostgres)
wrds <- dbConnect(Postgres(),
                  host='wrds-pgdata.wharton.upenn.edu',
                  port=9737,
                  dbname='wrds',
                  sslmode='require',
                  user='ct1424',
                  password = 'Pampers*9812')
```



## Extract datasets
```{r}
query1 = "
SELECT DISTINCT ON (DATE_TRUNC('month', date)) date,vix 
FROM cboe_all.cboe 
ORDER BY DATE_TRUNC('month', date), date DESC
"
query2 = "
SELECT date,spindx,sprtrn FROM crsp.msi
"
vix = dbGetQuery(wrds, query1)
sp500 = dbGetQuery(wrds, query2)

# Close the connection
dbDisconnect(wrds)

# Read csv file for 25 Portfolios sorted on Size and B/M
ports25 = read_csv("portfolios25_5x5.CSV",  skip = 15)

# Read csv file for Fama-French Factors
ff = read_csv("F-F_Research_Data_Factors.CSV",  skip = 3)
```

## Clean datasets
```{r}
# For vix
vix = vix %>%
  rename(vix_ind = 'vix') %>%
  mutate(
    vix_ind = ifelse(is.na(vix_ind), 0, vix_ind),
    date = ceiling_date(date, "month") - days(1)
    ) %>%
  filter(date >= as.Date("1990-01-31"))

vix = head(vix, -1)

# For sp500
sp500 = sp500 %>%
  mutate(
    date = ceiling_date(date, "month") - days(1) 
         ) %>%
  filter(date >= as.Date("1990-01-31"))

# For ff
names(ff)[1] <- "date"
ff = ff %>%
  mutate(
    date = as.Date(paste0(date, "01"), format = "%Y%m%d"),
    date = ceiling_date(date, "month") - days(1) 
         ) %>%
  filter(date >= as.Date("1990-01-31"))

# For ports25
names(ports25)[1] <- "date"
ports25 = ports25 %>%
  mutate(
    date = as.Date(paste0(date, "01"), format = "%Y%m%d"),
    date = ceiling_date(date, "month") - days(1) 
         ) %>%
  filter(date >= as.Date("1990-01-31")) %>%
  slice(1:420)
```

## Create main dataset with reg components
```{r}
reg_data = ports25 %>%
  pivot_longer(
    cols = -date,
    names_to = "port",
    values_to = "port_re"
  ) %>%
  arrange(port) %>%
  group_by(port) %>%
  left_join(ff, by = "date") %>%
  mutate(
    port_re = port_re - RF
  )
```

## Run Reg: (Step 1) TSR - i.e. across all time t, for each portfolio
```{r}
step1 = reg_data %>%
  nest(data = c(date,port_re,`Mkt-RF`,SMB,HML,RF)) %>%
  mutate(b_estimates = map(
    data,
    ~tidy(lm(port_re ~ `Mkt-RF` + SMB + HML, data = .x))
  )) %>%
  unnest(b_estimates) %>%
  select(port, estimate, term) %>%
  pivot_wider(names_from = term,
              values_from = estimate) %>%
  select(port,
         b_mkt = "`Mkt-RF`",
         b_smb = SMB,
         b_hml = HML)

print(step1)

step1 = reg_data %>%
  left_join(step1, by = "port")
```

## Run Reg: (Step 2) CSR - i.e. across the diff portfolios, at each time t
```{r}
step2 = step1 %>%
  nest(data = c(port,port_re,b_mkt,b_smb,b_hml)) %>%
  mutate(l_estimates = map(
    data,
    ~tidy(lm(port_re ~ b_mkt + b_smb + b_hml, data = .x))
  )) %>%
  unnest(l_estimates) %>%
  select(date, estimate, term) %>%
  pivot_wider(names_from = term,
              values_from = estimate) %>%
  select(date,
         l_mkt = b_mkt,
         l_smb = b_smb,
         l_hml = b_hml)

print(step2)
```

## Find factor risk premiums: (Step 3): Calc mean and var/standard error of lambdas
```{r}
l_mkt_mean = mean(step2$l_mkt)
l_smb_mean = mean(step2$l_smb)
l_hml_mean = mean(step2$l_hml)

l_mkt_var = var(step2$l_mkt)
l_smb_var = var(step2$l_smb)
l_hml_var = var(step2$l_hml)

```

## Estimate cross-sectional averages
```{r}
t.test(step1$b_mkt, mu=0)
t.test(step1$b_smb, mu=0)
t.test(step1$b_hml, mu=0)

t.test(step2$l_mkt, mu=0)
t.test(step2$l_smb, mu=0)
t.test(step2$l_hml, mu=0)
```

######
## Augmenting reg with first log diff of VIX
```{r}
vix = vix %>%
  #vix_ind = log(vix_ind) %>%
  mutate(
    vix_lg_diff = vix_ind - lag(vix_ind)
  )

areg_data = ports25 %>%
  pivot_longer(
    cols = -date,
    names_to = "port",
    values_to = "port_re"
  ) %>%
  arrange(port) %>%
  group_by(port) %>%
  left_join(ff, by = "date") %>%
  left_join(vix, by = "date") %>%
  ungroup() %>%
  mutate(
    port_re = port_re - RF
  ) %>%
  na.omit()
```

## Run Reg: (Step 1) TSR - i.e. across all time t, for each portfolio
```{r}
astep1 = areg_data %>%
  nest(data = c(date,port_re,`Mkt-RF`,SMB,HML,RF, vix_ind, vix_lg_diff)) %>%
  mutate(b_estimates = map(
    data,
    ~tidy(lm(port_re ~ `Mkt-RF` + SMB + HML + vix_lg_diff, data = .x))
  )) %>%
  unnest(b_estimates) %>%
  select(port, estimate, term) %>%
  pivot_wider(names_from = term,
              values_from = estimate) %>%
  select(port,
         b_mkt = "`Mkt-RF`",
         b_smb = SMB,
         b_hml = HML,
         b_vld = vix_lg_diff)

print(astep1)

astep1 = areg_data %>%
  left_join(astep1, by = "port")
```

## Run Reg: (Step 2) CSR - i.e. across the diff portfolios, at each time t
```{r}
astep2 = astep1 %>%
  nest(data = c(port,port_re,b_mkt,b_smb,b_hml, b_vld)) %>%
  mutate(l_estimates = map(
    data,
    ~tidy(lm(port_re ~ b_mkt + b_smb + b_hml + b_vld, data = .x))
  )) %>%
  unnest(l_estimates) %>%
  select(date, estimate, term) %>%
  pivot_wider(names_from = term,
              values_from = estimate) %>%
  select(date,
         l_mkt = b_mkt,
         l_smb = b_smb,
         l_hml = b_hml,
         l_vld = b_vld)

print(astep2)
```

## Find factor risk premiums: (Step 3): Calc mean and var/standard error of lambdas
```{r}
al_mkt_mean = mean(astep2$l_mkt)
al_smb_mean = mean(astep2$l_smb)
al_hml_mean = mean(astep2$l_hml)
al_vld_mean = mean(astep2$l_vld)

al_mkt_var = var(astep2$l_mkt)
al_smb_var = var(astep2$l_smb)
al_hml_var = var(astep2$l_hml)
al_vld_var = var(astep2$l_vld)
```

## Estimate cross-sectional averages
```{r}
t.test(astep1$b_mkt, mu=0)
t.test(astep1$b_smb, mu=0)
t.test(step1$b_hml, mu=0)

t.test(astep2$l_mkt, mu=0)
t.test(astep2$l_smb, mu=0)
t.test(astep2$l_hml, mu=0)
t.test(astep2$l_vld, mu=0)
```












###### Archive

## Run Reg: (Step 1) TSR - i.e. across all time t, for each portfolio
```{r}

# Get estimated betas - for each portfolio
b_est = data.frame(matrix(ncol = 4, nrow = 25)) %>%
  rename(
    intercept = 'X1',
    b1 = 'X2',
    b2 = 'X3',
    b3 = 'X4'
    )
col_names = data.frame(colnames(ports25)[2:ncol(ports25)])

for (i in 1:nrow(col_names)){
  port = col_names[i,]
  y = as.matrix(ports25[port])  # Dependent variable (portfolio i's excess returns)
  x = as.matrix(reg_data[,(4:6)])  # independent variable (fama factors)
  
  mod = lm(y ~ x)
  b = mod$coefficients
  b[is.na(mod$coefficients)] = 0
  
  b_est[i,] = b
}
```

```{r}

# Get estimated lambdas - for each time t
l_est = data.frame(matrix(ncol = 4, nrow = nrow(ports25)*25)) %>%
  rename(
    mkt = 'X1',
    smb = 'X2',
    hml = 'X3',
    port = 'X4'
    )

col_names = data.frame(colnames(ports25)[2:ncol(ports25)])

for (i in 2:(nrow(col_names)+1)){
  
  for (t in 1:nrow(ports25)){
  
    y = as.matrix(ports25[t,i])  # Portfolio i's excess return at a t)
    x = as.matrix(b_est[i-1,(2:4)])  # est betas
  
    mod = lm(y ~ x - 1)
    l = mod$coefficients
    l[is.na(mod$coefficients)] = 0
  
    l_est[t,1] = l[1]
    l_est[t,2] = l[2]
    l_est[t,3] = l[3]
    l_est[t,4] = col_names[i-1,1]
  }
}

l_est$mkt = mean(ff$`Mkt-RF`)
l_est$smb = mean(ff$`SMB`)
l_est$hml = mean(ff$`HML`)

```








